---
title: "ISDS Mid-term Exam"
author: "Tyler McDonald"
date: "September 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
load(file = 'employeesBR.RData')
dt <- dt %>% left_join(lkt)
```

## Introduction

  In the dataset **dt** which has been joined with dataset **lkt**, there are *`r length(dt)`* rows.
  
## Data Inspection

  There are a total of `r count(distinct(dt[3]))` departments in this dataset.

```{r}
employee_count <- dt %>% group_by(department_name) %>% count(department_name) %>% arrange(desc(n))
employee_count
```

  The departments in order of number of employees are as follows: `r employee_count$department_name`. The total number of observations in MOSQUITO & RODENT CONTROL IS `r employee_count$n[25]`.

## Data Preparation

```{r}
dt <- dt %>% filter(department_name != 'COUNCIL ADMINISTRATOR')
retire <- dt %>% mutate(retirement = if_else(is.na(end_date), FALSE, TRUE)) %>% group_by(department_name) %>% count(retirement)
retir <- arrange(retire,retirement,desc(n))
```

```{r}
retir %>% group_by(department_name) %>% mutate(total = sum(n)) %>% filter(retirement == TRUE) %>% mutate(percent = n/total)
```

```{r}
retir2 <- retir %>% group_by(retirement) %>% summarise(total = sum(n))
retir2
```

A total of `r retir2$total[2]` people retired while `r retir2$total[1]` did not.

## Data Visualization

```{r}
ggplot() + geom_bar(aes(x = retir$department_name, weight = retire$n)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
avg_base_pay <- dt %>% group_by(department_name) %>% summarise(average = mean(base_pay))
ggplot(data = avg_base_pay) + geom_bar(aes(x = department_name, weight = average)) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
avg_base_pay
```

The department with the largest base pay is `r filter(avg_base_pay, average == max(average))`.

```{r}
avg_base_pay %>% arrange(desc(average)) %>% slice(1:10)
```

```{r}
dt %>% mutate(costOvertime = overtime_hourly_rate * total_overtime_hours) %>% group_by(department_name) %>% summarise(average = mean(costOvertime)) %>% arrange(desc(average)) %>% slice(1:3)
```