---
title: "Final Exam"
author: "Name:"
output:
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, fig.width = 4, fig.height = 3)
```

\vspace{30pt}

For each question, change the code to produce the **correct** output. Each question contains only one mistake, but which may be repeated (e.g., using the wrong function multiple times)

The following libraries are loaded in the global environment when running each snippet of code:

```{r}
library(tidyverse)
library(fivethirtyeight)
```

\vspace{30pt}

\newpage

## Data Visualization

1.

\vspace{30pt}

```{r, error=TRUE}
ggplot(data = steak_survey) +
  geom_col(aes(x = region)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**correct:**

```{r, echo=FALSE}
#correct
ggplot(data = steak_survey) +
  geom_bar(aes(x = region)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


\newpage

2. Count of subjects in each level of steak preparation preference by gender (using variable `female`) 

\vspace{30pt}

```{r}
ggplot(data = steak_survey) +
  geom_bar(aes(x = steak_prep, fill = 'darkblue')) +
  facet_wrap(~ lottery_a) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
#correct
ggplot(data = steak_survey) +
  geom_bar(aes(x = steak_prep), fill = 'darkblue') +
  facet_wrap(~ lottery_a) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

\vspace{30pt}

\newpage

3. Chart of the correlation between `insurance_premiums` and `losses`

\vspace{30pt}

```{r, error=TRUE}
ggplot(bad_drivers) +
  geom_smooth(aes(insurance_premiums, losses), method = 'lm') +
  geom_point()
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
ggplot(bad_drivers, aes(insurance_premiums, losses)) +
  geom_smooth(method = 'lm') +
  geom_point()
```

\newpage

# Data Analysis

4. Total graduates by major_category (first 5 rows)

```{r, }
college_recent_grads %>% 
  group_by(major_category) %>% 
  summarise(Total = sum(total)) %>% 
  slice(1:5)
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
#correct
college_recent_grads %>% 
  filter(!is.na(total)) %>% 
  group_by(major_category) %>% 
  summarise(Total = sum(total)) %>% 
  slice(1:5)
```

\vspace{30pt}

5. Total share of women by major_category (first 5 rows)

```{r, error=TRUE}
college_recent_grads %>% 
  group_by(major_category) %>% 
  summarise(percent = n(women)/n(total)) %>% 
  slice(1:5)
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
college_recent_grads %>% 
  group_by(major_category) %>% 
  summarise(percent = sum(women)/sum(total)) %>% 
  slice(1:5)
```

\newpage

6. Extract the row for the major with the highest share of females

\vspace{30pt}

```{r, error=TRUE}
college_recent_grads %>% 
  filter(max(sharewomen, na.rm = T)) %>% 
  select(major_category:sharewomen)
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
college_recent_grads %>% 
  filter(sharewomen == max(sharewomen, na.rm = T)) %>% 
  select(major_category:sharewomen)
```

\vspace{30pt}

7. Count of major_category for Engineering and Business only.

\vspace{30pt}
```{r, error=TRUE}
college_recent_grads %>% 
  filter(major_category == 'Engineering', major_category == 'Business') %>% 
  count(major_category) 
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
college_recent_grads %>% 
  filter(major_category == 'Engineering' | major_category == 'Business'  ) %>% 
  count(major_category) 
```

\newpage


8. Transform `shareswomen` in percent. Print only the first row and keep only column "sharewoman"

\vspace{30pt}

```{r, error=TRUE}
college_recent_grads %>% 
  slice(1) %>% 
  select(sharewomen) %>% 
  filter(sharewomen = scales::percent(sharewomen))
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
college_recent_grads %>% 
  slice(1) %>% 
  select(sharewomen) %>% 
  mutate(sharewomen = scales::percent(sharewomen))
```
 
\vspace{30pt}

9. Output a vector of means

```{r, error=TRUE}
map_dbl(mean, mtcars)
```

**correct:**

```{r, echo=FALSE}
map_dbl(mtcars, mean)
```

\newpage 

# Functions

10. Calculate the mean for each column of a dataframe containing numeric values:

```{r, error=TRUE}
columnwiseMean <- function(df) {
  if (is.data.frame(df)) stop('`df` is not a dataframe')
  map(df, mean)
} 
columnwiseMean(cars)
```

\vspace{30pt}
**correct:**


```{r, echo= FALSE}
columnwiseMean <- function(df) {
  #either delete the if-statement or  !is.data.frame()
  map(df, mean)
} 
columnwiseMean(cars)
```


11. Calculate a percentage given a total amount and a fraction of the total observations


```{r, error=T}
prc <- function(total, obs, call. = FALSE) {
          if (!is.double(total) | !is.double(obs)) {
            stop('`total` and `obs` must be of type numeric')  
            }
         (obs / total) * 100
         x <- paste0(x, '%')
         x
}
prc(400, 20)
```

\vspace{30pt}
**correct:**

```{r, echo=FALSE}
prc <- function(total, obs, call. = FALSE) {
          if (!is.double(total) | !is.double(obs)) {
            stop('`total` and `obs` must be of type numeric')  
            }
         x <- (obs / total) * 100
         x <- paste0(x, '%')
         x
}
```
```{r}
prc(400, 20)
```



\newpage

#General R knowledge (TRUE/FALSE)


  - `==` performs an assignment operation : F
 
  - a `NA` is equivalent to `0` : F
 
  - In dataframes, each column must have the same number of rows/cells: T
 
  - The code snippets below retrieve the same output: T (vectorization)
 
```{r, eval=FALSE}
c(1, 4, 3) + 1           # 1.

c(1, 4, 3) + c(1, 1, 1)  # 2. 
```
 

  - The code snippets below retrieve the same output when `x` is the same: F
 
```{r, eval=FALSE}
mean(is.na(x))             # 1. 

x %>% mean() %>% is.na(x)  # 2.
```
 

  - The code snippets below retrieve the same output when `x` is the same: T
 
```{r, eval=FALSE}
bad_drivers %>% filter(num_drivers > x)    # 1. 

bad_drivers %>% filter(!(num_drivers <= x))  # 2.
```



  - The code snippets below retrieve the same output: F
 
```{r, eval=FALSE}
steak_survey %>% count(lottery_a, smoke)          # 1. 

steak_survey %>% group_by(lottery_a, smoke) %>%   # 2.
  summarise(sum())
```


  - `filter` is for manipulating rows, while `slice` is for manipulating variables: F
  

  - the output from `summarise()` has as many rows as the grouping levels: T

  
  - datasets being plotted in `ggplot()` should be "tidy" (or in longitudinal form): T
  
  
  - in ggplot, we can map multiple variables to the same aesthetic: F
  
  
  - every time we query a database using `dplyr` the output is automatically collected locally on our machine: F

