---
title: "lesson 8: Querying a MySQL DB"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=T)
library(tidyverse)
source(here::here('lectures/lesson08_notes_datamodel/conn.R'))
```

#. Report the top 3 hotels with the best service (calculate the average using review.reviewService)

```{r}
tbl(con, 'hotel') %>% inner_join(tbl(con, 'review'), by = c('hotelId' = 'hotelId')) %>% 
  group_by(hotelName) %>% 
  summarise(avg = mean(reviewService)) %>% 
  arrange(desc(avg)) %>% 
  collect() %>% slice(1:3)
```


#. Report all the authors and the hotels they reviewed. Report the author name, the hotel ID and hotel name. Order the output by author name

```{r}
tbl(con, 'hotel') %>%
 inner_join(tbl(con, 'review')) %>%
 inner_join(tbl(con, 'author')) %>%
 select(authorNickname, hotelName, hotelId) %>%
 arrange(desc(authorNickname))

```


#. Report the total number of reviews received by each hotel. Report a three columns table with `hotelId`, `hotelName` and total number of reviews. Make sure that you are including hotels with zero reviews as well. Note that some hotels have 0 reviews. If you decide to count the rows in each grouping level, make sure you you don't count as 1 those who have zero reviews.

```{r}
dt <- tbl(con, sql('SELECT hotel.hotelId, hotel.hotelName, COUNT(reviewId) FROM hotel, review WHERE hotel.hotelId = review.hotelId GROUP BY review.hotelId'))

tbl(con, 'hotel') %>%
 left_join(tbl(con, 'review')) %>%
  select(hotelId, hotelName, reviewId) %>% 
  count(hotelId, hotelName, wt = !is.na(reviewId), sort = T)
```


#. Report how many chains established their headquarter in each country. Plot a barchart those frequencies by country and then add a title to your plot. Adjust your axes' labels as needed (remember that non-data ink points are modified within `theme()`).

```{r}
tbl(con, 'lodgingchain') %>% 
  count(headquarters) %>% 
  ggplot() +
  geom_col(aes(x = headquarters, y = n)) +
  theme(axis.text.x = element_text(angle  = 45, hjust = 1))
```


#. For each review of stays during 2011, calculate the composite average score as the mean of Location, Room, Service, Value and Cleanliness. Then plot a histogram (using `geom_hist()`) of the differences between the average of those 5 attributes and ratingOverall.

```{r}
tbl(con, 'review') %>% filter(reviewStayYear == 2011) %>% 
  collect() %>% 
  gather('attribute', 'score', reviewCleanliness:reviewValue) %>% 
  group_by(reviewId) %>% summarise(compAvg = mean(score), overall = mean(reviewOverall)) %>% 
  mutate(diff = compAvg - overall) %>% 
  ggplot() +
  geom_histogram(aes(x = diff))
```

*warning*: If you query the DB and then pipe the remote table into a ggplot2 function, you might need to `collect()` the remote table before passing it to `ggplot()`, because `ggplot()` expects a class `data.frame` (and tibbles are also data.frame). However, if you pass a different object (such as a connection) "it will be converted to one by fortify()" (see the documentation `?ggplot()`). 
Apparently, for some of you `fortify()` does not convert the table to a local data.frame, and if that is the case you need to `collect()`.


#. Report all data about brands that have minimum square footage requirement for rooms that exceeds the average minimum square footage of all brands by at least 50%.

```{r}
tbl(con, 'brand') %>% 
  filter(!is.na(brandMinSqFt)) %>% collect() %>% 
  filter(brandMinSqFt >= mean(brandMinSqFt)*1.5)
```




